Efficient Coalition Formation \\ for Web Services

### 2013-07-27 papers/coalition-journal-extension/coalition_journal.tex
Changes

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed mechanisms to help
users and services maximize their gain. A two-player
non-cooperative game between web services and community master was
introduced in \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this
game-theoretic model, the strategies available to a web service
when facing a new community are requesting to join the community,
accepting the master's invitation to join the community, or
refusing the invitation to join. The set of strategies for
communities are inviting the web service or refusing the web
service's join request. Based on their capacity, market share and
reputation, the two players have different set of utilities over
the strategy profiles of the game. The main limits of this game
model are: 1) its consideration of only three quality parameters,
while the other factors are simply ignored; and 2) the
non-consideration of the web services already residing within the
community. The game is only between the community master and the
new web service, and the inputs from all the other members are
simply ignored. The consideration of those inputs is a significant
issue as existing web services can lose utility or payoff because
of the new member, which can results in an unhealthy and unstable
group. The problem comes from the fact that the existing members
should collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

In \cite{DBLP:conf/IEEEscc/LimTMB12} a 3-way satisfaction approach
for selecting web services has been proposed. In this approach,
the authors proposed a web service selection process that the
community masters can use. The approach considerers the efficiency
of all the three involved parties, namely users, web services and
communities. In this work, it is shown how the gains of these
parties are coupled together using a linear optimization process.
However, the optimization problem in this solution tends to
optimize some parameters considering all web services regardless
of their efficiency and contribution to the community's welfare.
Moreover, there are no clear thresholds for accepting or rejecting
new web services. The solution of the optimization problem could,
for instance, suggest web services already residing within the
community to increase or decrease their capacity to cover up the
weakness of other parties in the system. However, a high
performing web service could deviate anytime it finds itself
unsatisfied within the community instead of adjusting its service
parameters.

In \cite{10.1109/TSC.2012.12} a cooperative scheme among autonomous
web services based on coalitional game theory has been introduced. They have proposed an algorithm to
reach individually stable coalition partition for web services in order to
maximize their efficiency. The communities choose new web services on the promise
that it would benefit the community without decreasing any other web service's
income. In their model, the worth of community is evaluated with high emphasis on
availability metric and considering price and cost values only. The community structure is based on a coordination chain,
where a web service is assigned as a \emph{primary} web service and the community task destribution
method, will initially invoke the primary web service and only if the primary web service is unavailable
will invoke the next backup web services as they are ordered in the coordination chain. However in cooperative models, it is preferred to
have a real and active cooperative activity engaging all agents to perform the tasks more efficiently. Especially nowadays
with recent advancement in cloud and hardware infrastructures availability is becoming less of an issue. So the backup web services
in their model have a very low chance of getting jobs, especially the ones further in chain, which is huge waste of web services
capabilities.

@@

\emph{1) Web services} are rational entities\footnote{The term
rational is used here in the sense that web services are utility
maximizers} providing services to end users. They aim to maximize
their individual income by receiving enough requests from end
users. In order to increase their revenue, web services seek for
more tasks if they have the capacity and throughput to do so. Web
services can join communities to have better efficiency by
collaborating with others, to have access to higher market share,
and to have opportunity of receiving a bigger task pool from end
users. Throughout this paper, in
our equations, we refer to web services as $ws$ and to the set of
web services hosted by a given community as $C$. To simply the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

@@

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players. Any
subset $S$ of $N$ can form a group called $coalition$. A
\emph{coalitional game} is a pair $G = (N, v)$ where $v$ is called
a \emph{characteristic function}, which given a coalition, is a
function $v: 2^N \to \mathbb{R}$ mapping the set of players of the
coalition to a real number $v(S)$, the worth of $S$. This number
usually represents the output or payoff or again the performance
of these players working together as coalition.  If a coalition
$S$ is formed, then it can divide its worth, $v(S)$ in any
possible way among its members. The payoff vector $x \in
\mathbb{R}^S$ is the amount of payoff being distributed among the
members of the coalition $S$. The payoff vector satisfies two
conditions:

@@

The second criteria is called the \emph{feasibility} condition,
according to which, the payoff for each agent cannot be more than
the coalition total gain. A payoff vector is also \emph{efficient}
if the payoff obtained by a coalition is distributed amongst the
coalition members: $\sum_{i \in S} x_i = v(S)$. This definition of
the characteristic function works in \emph{transferable utility}
(TU) settings, where utility (i.e., payoff) is transferable from
one player to another, or in other words, players have common
currency and a unit of income is worth same for all players
\cite{myerson1991game}.

When dealing with cooperative games, two issues need to be
addressed:\\ 1. What coalitions to form? \\
2. How to reward each member when a task is completed?\\
%
The following definitions help address these two issues.

@@

The communities and web services earn revenue by performing tasks. The total gain is function of quality ($QoS_{ws}$) and quantity($Tr_{ws}$) of tasks being performed. As mentioned in section \ref{ws_parameters}, $QoS_{ws}$ is obtained by a certain aggregation function of the parameters considered in Table \ref{qosws}. We have adopted a linear equal weigth average over all QoS parameteres listed in table \ref{qosws} excluding the $Throuput$ and $Cost$ parameters. A community has the option to weigh specific QoS patameters more or less depending on the expectations of their clients.

@@

We present an example with a community with $R_C$ value of 10, and three web services each having different $QoS_{ws}$ and $Th_{ws}$ values as listed in table \ref{example_1}. The worth of a community is calculated based on $Out(C)$ equation (\ref{out_c}) which is the amount of output being generated by the community. The first table in figure \ref{example_1} lists the web services with their aggregated $QoS_{ws}$ parameters, their task input rate while working alone, and also their $Th_{ws}$ throughput value. The second table shows all the possible communities and their worth. The values suggest, communities having more web services have better gain and output. However community tries to distribute the gain between web services. Sometimes it is impossible to share the gain between all web services in a way no subset of them would individually gain more if they had their own group. In this example, the value community of ${ws_1}$ and ${ws_2}$ is 7.2, with ${ws_3}$ joining the community the value of community increases to 8.0, however there is no way to distribute the value among web services, where ${ws_1}$ and ${ws_2}$ can earn 7.2, and ${ws_3}$ earn at least 2.4, which they could earn before joining the community. This makes the group unstable. In the example of table \ref{example_2}, we even have situations where a web service (${ws_3}$) joining a community ($\left\{ws_1,ws_2\right\}$) decreases the value of community. The reason is, the community is already full and all tasks are almost being distributed and new community with bad quality can degrade the average quality of tasks being done by the community. In both examples, the request of joining of web service ${ws_3}$ should be rejected by the community.

@@

Now we consider another example, in which we have different communities with different market share, ${R_C}$ values. Web services also have a small share of market independently, providing them with a small task pull. In these kind of scenarios, the solution considers individual maximization of payoff and also the total worth of all communities which represents the $social welfare$. In this example the most efficient partition of web services is earned by having two coalitions of $\left\{C_1, ws_2\right\}$ and $\left\{C_2, ws_1, ws_3\right\}$, which yields a total value of $32 + 16 = 48$. In these types of scenarios, the goal is to reach stability, adopting a distributed approach where all players have the power of choice on the decision of whether or not they join a coalition. The communities usually start the game having some established members, encountering new web services, the communities may exchange web services and new web services would join them having at least one player gaining utility, without hurting any other participant. In this example if we initially having two coalitions of $\left\{C_1, ws_2\right\}$ and $\left\{C_2, ws_1\right\}$ and a ${ws_3}$ as new web service, ${ws_3}$ joining ${C_1}$ would hurt at least itself or $ws_2$, however ${ws_3}$ joining ${C_2}$ would not hurt any participants and ${ws_3}$ would earn more within the community and the community will have enough web services performing the incoming tasks from users.

@@

We discuessed $core$ as one the preminent solution concepts in cooperative games. Working together, completing tasks and generating revenue, agents need to distribute the gain in a way no agents would gain more by having their own group. However, in most cases the core of a game is empty so we introduced the \emph{$\epsilon$-core} concept, where agents would only earn a minimal ammount of $\epsilon$ by deviating. Comuunity stability is an attractice property for communities, also communities would benefit by having slightly more web services then having the exact number of web services satisfying task rate cap. One reason is there is always possibility that agents may leave the community or they may under perform from the quality values they were initially performing with.

One way for communities to ensure community stability could be applying a tax, $\epsilon$ ammount of cost for web services changing communities which would make deviation a costly act. However this would require all community coordinators to agree on same ammount of taxtation, being governed by some external entities or web services would join communities with lesser ammount of tax. Another viable solution in our scenario is to stabilize the game using external subsidies. The reason a game is not stable is that the community is not making enough revenue to allocate the players with enough gain. A community master can subside its community with a constant coefficient value of $\lambda$. Rewarding a community with high number of quality participants with $\lambda v(C)$, where $\lambda \leq 1$. Obviously with a big number of $\lambda$ it is always possible to stabilize the community. However, this can be a costly act for community masters so they are interested in the minimum subside value of $\lambda$ making the game stable. Subsiding or taxting in order to reduce the bargaining power of sub coalitions are called $taxtation$\cite{eps346856} methods.

>>>>

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed mechanisms to help
users and services to maximize their gain. A two-player
non-cooperative game between web services and community master was
introduced in \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this
game-theoretic model, the strategies available to a web service
when facing a new community are requesting to join the community,
accepting the master's invitation to join the community, or
refusing the invitation to join. The set of strategies for
communities are inviting the web service or refusing the web
service's join request. Based on their capacity, market share and
reputation, the two players have different set of utilities over
the strategy profiles of the game. The main limits of this game
model are: 1) its consideration of only three quality parameters,
while the other factors are simply ignored; and 2) the
non-consideration of the web services already residing within the
community. The game is only between the community master and the
new web service, and the inputs from all the other members are
simply ignored. The consideration of those inputs is a significant
issue as existing web services can lose utility or payoff because
of the new member, which can results in an unhealthy and unstable
group. The problem comes from the fact that the existing members
should collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

In \cite{DBLP:conf/IEEEscc/LimTMB12}, a 3-way satisfaction approach
for selecting web services has been proposed. In this approach,
the authors proposed a web service selection process that the
community masters can use. The approach considers the efficiency
of all the three involved parties, namely users, web services and
communities. In this work, it is shown how the gains of these
parties are coupled together using a linear optimization process.
However, the optimization problem in this solution tends to
optimize some parameters considering all web services regardless
of their efficiency and contribution to the community's welfare.
Moreover, there are no clear thresholds for accepting or rejecting
new web services. The solution of the optimization problem could,
for instance, suggest web services already residing within the
community to increase or decrease their capacity to cover up the
weakness of other parties in the system. However, a high
performing web service could deviate anytime it finds itself
unsatisfied within the community instead of adjusting its service
parameters.

In \cite{10.1109/TSC.2012.12}, a cooperative scheme among autonomous
web services based on coalitional game theory has been introduced. They have proposed an algorithm to
reach individually stable coalition partition for web services in order to
maximize their efficiency. The communities choose new web services on the promise
that it would benefit the community without decreasing any other web service's
income. In their model, the worth of community is evaluated with high emphasis on
availability metric and considering price and cost values only. The community structure is based on a coordination chain,
where a web service is assigned as a \emph{primary} web service and the community task destribution
method, will initially invoke the primary web service and only if the primary web service is unavailable
will invoke the next backup web services as they are ordered in the coordination chain. However in cooperative models, it is preferred to
have a real and active cooperative activity engaging all agents to perform the tasks more efficiently. Especially nowadays
with recent advancement in cloud and hardware infrastructures availability is becoming less of an issue. So the backup web services
in their model have a very low chance of getting jobs, especially the ones further in chain, which is huge waste of web services
capabilities.

@@

\emph{1) Web services} are rational entities\footnote{The term
rational is used here in the sense that web services are utility
maximizers} providing services to end users. They aim to maximize
their individual income by receiving enough requests from end
users. In order to increase their revenue, web services seek for
more tasks if they have the capacity and throughput to do so. Web
services can join communities to have better efficiency by
collaborating with others, to have access to higher market share,
and to have opportunity of receiving a bigger task pool from end
users. Throughout this paper, in
our equations, we refer to web services as $ws$ and to the set of
web services hosted by a given community as $C$. To simplify the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

@@

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players. Any
subset $S$ of $N$ can form a group called $coalition$. A
\emph{coalitional game} is a pair $G = (N, v)$, where $v$ is called
a \emph{characteristic function} $v: 2^N \to \mathbb{R}$, mapping the set of players of the
coalition to a real number $v(S)$, the worth of $S$. This number
usually represents the output or payoff or again the performance
of these players working together as coalition.  If a coalition
$S$ is formed, then it can divide its worth, $v(S)$ in any
possible way among its members. The payoff vector $x \in
\mathbb{R}^S$ is the amount of payoff being distributed among the
members of the coalition $S$. The payoff vector satisfies two
conditions:

@@

The second criteria is called the \emph{feasibility} condition,
according to which, the payoff for each agent cannot be more than
the coalition total gain. A payoff vector is also \emph{efficient}
if the payoff obtained by a coalition is distributed amongst the
coalition members: $\sum_{i \in S} x_i = v(S)$. This definition of
the characteristic function works in \emph{transferable utility}
(TU) settings, where utility (i.e., payoff) is transferable from
one player to another, or in other words, players have common
currency and a unit of income that is worth the same for all players
\cite{myerson1991game}.

When dealing with cooperative games, two issues need to be
addressed:\\ 1. Which coalitions among all possible coalitions to form? \\
2. How to reward each member when a task is completed?\\
%
The following definitions help address these two issues.

@@

The communities and web services earn revenue by performing tasks. The total gain is function of quality ($QoS_{ws}$) and quantity ($Tr_{ws}$) of tasks being performed. As mentioned in section \ref{ws_parameters}, $QoS_{ws}$ is obtained by a certain aggregation function of the parameters considered in Table \ref{qosws}. We have adopted a linear equal weigth average over all QoS parameteres listed in table \ref{qosws} excluding the $Throuput$ and $Cost$ parameters. A community has the option to weight specific QoS patameters more or less depending on the expectations of their clients.

@@

We present an example with a community with $R_C$ value of 10, and three web services each having different $QoS_{ws}$ and $Th_{ws}$ values as listed in table \ref{example_1}. The worth of a community is calculated based on $Out(C)$ equation (\ref{out_c}) which is the amount of output being generated by the community. The first table in figure \ref{example_1} lists the web services with their aggregated $QoS_{ws}$ parameters, their task input rate while working alone, and also their $Th_{ws}$ throughput value. The second table shows all the possible communities and their worth. The values suggest, communities having more web services have better gain and output. However community tries to distribute the gain between web services. Sometimes it is impossible to share the gain between all web services in a way no subset of them would individually gain more if they had their own group. In this example, the value community of ${ws_1}$ and ${ws_2}$ is 7.2, with ${ws_3}$ joining the community the value of community increases to 8.0. However there is no way to distribute the value among web services, where ${ws_1}$ and ${ws_2}$ can earn 7.2, and ${ws_3}$ earn at least 2.4, which they could earn before joining the community. This makes the group unstable. In the example of table \ref{example_2}, we even have situations where a web service (${ws_3}$) joining a community ($\left\{ws_1,ws_2\right\}$) decreases the value of community. The reason is, the community is already full and all tasks are almost being distributed and new community with bad quality can degrade the average quality of tasks being done by the community. In both examples, the request of joining of web service ${ws_3}$ should be rejected by the community.

@@

Now we consider another example, in which we have different communities with different market share, ${R_C}$ values. Web services also have a small share of market independently, providing them with a small task pull. In these kind of scenarios, the solution considers individual maximization of payoff and also the total worth of all communities which represents the \emph{social welfare}. In this example the most efficient partition of web services is earned by having two coalitions of $\left\{C_1, ws_2\right\}$ and $\left\{C_2, ws_1, ws_3\right\}$, which yields a total value of $32 + 16 = 48$. In these types of scenarios, the goal is to reach stability, adopting a distributed approach where all players have the power of choice on the decision of whether or not they join a coalition. The communities usually start the game having some established members, encountering new web services, the communities may exchange web services and new web services would join them having at least one player gaining utility, without hurting any other participant. In this example if we initially having two coalitions of $\left\{C_1, ws_2\right\}$ and $\left\{C_2, ws_1\right\}$ and a ${ws_3}$ as new web service, ${ws_3}$ joining ${C_1}$ would hurt at least itself or $ws_2$, however ${ws_3}$ joining ${C_2}$ would not hurt any participants and ${ws_3}$ would earn more within the community and the community will have enough web services performing the incoming tasks from users.

The first two examples illustrate the fact that a community cannot simply increase its revenue by adding more web services. The web services and even community owners are autonomous agents and would deviate and be displeased about the community if new members cause a drop in their profit. The job of the community master is to attract as many quality web services it can and keep them satisfied; hence the group stability is guaranteed. 
The third example highlights another type of problem we would like to address, which is how to form best possible groups of communities, and allocate web services among communities in a way which would maximize payoff for of our agents and members already residing in the communities.
In next section, we provide collaborative game theory based algorithms for our autonomous agents, to tackle these problems and find applicable and efficient strategies for communities and web services to maximize their profit.

@@

We discuessed $core$ as one of the preminent solution concepts in cooperative games. Working together, completing tasks and generating revenue, agents need to distribute the gain in a way no agents would gain more by having their own group. However, in most cases the core of a game is empty so we introduced the \emph{$\epsilon$-core} concept, where agents would only earn a minimal ammount of $\epsilon$ by deviating. Comuunity stability is an attractice property for communities, also communities would benefit by having slightly more web services then having the exact number of web services satisfying task rate cap. One reason is that there is always the possibility that the agents may leave the community or they may under perform from the quality values they were initially performing with.

One way for communities to ensure community stability could be applying a tax, $\epsilon$ ammount of cost for web services changing communities which would make deviation a costly act. However this would require all community coordinators to agree on same ammount of taxtation, being governed by some external entities or web services would join communities with lesser ammount of tax. Another viable solution in our scenario is to stabilize the game using external subsidies. The reason a game is not stable is that the community is not making enough revenue to allocate the players with enough gain. A community master can subside its community with a constant coefficient value of $\lambda$. Rewarding a community with high number of quality participants with $\lambda v(C)$, where $\lambda \leq 1$. Obviously with a big number of $\lambda$ it is always possible to stabilize the community. However, this can be a costly act for community masters so they are interested in the minimum subside value of $\lambda$ making the game stable. Subsiding or taxting in order to reduce the bargaining power of sub coalitions are called $taxtation$ \cite{eps346856} methods.
